Gene expression responses to diet quality and viral infection in Apis mellifera

### 2018-09-09 Manuscript/Text/manuscript.tex
Ethics statement
Tukey versus Bonferroni

As shown in Figure 1, mortality rates of honey bees 72 hour post-inoculation significantly differed among the treatment groups (mixed model ANOVA across all treatment groups, df = 3, 54; F = 10.03; p $<$ 2.30e-05). The effect of virus treatment (mixed model ANOVA, df = 1, 54; F = 24.73; p $<$ 1.00e-05) and diet treatment (mixed model ANOVA, df = 1, 54; F = 5.32; p $<$ 2.49e-02) were significant, but the interaction between the two factors (mixed model ANOVA, df = 1, 54; F = 4.72e-02, p = 8.29e-01) was not significant. We compared mortality levels based on pairwise comparisons: For a given diet, honey bees exposed to the virus showed significantly higher mortality rate than honey bees not exposed to the virus. Namely, bees fed Rockrose pollen had significantly elevated mortality with virus infection compared to uninfected controls (Tukey HSD, p $<$ 1.18e-03), and bees fed Chestnut pollen similarly had significantly elevated mortality with virus infection compared to controls (Tukey HSD, p $<$ 4.80e-03) (Figure 1).

As shown in Figure 2, IAPV titers of honey bees 72 hour post-inoculation significantly differed among the treatment groups (mixed model ANOVA across all treatment groups, df = 3, 33; F = 6.10; p $<$ 1.96e-03). The effect of virus treatment (mixed model ANOVA, df = 1, 33; F = 15.04; p $<$ 4.75e-04) was significant, but the diet treatment (mixed model ANOVA, df = 1, 33; F = 2.55; p = 1.20e-01) and the interaction between the two factors (mixed model ANOVA, df = 1, 33; F = 7.02e-01, p = 4.08e-01) were not significant. We compared IAPV titer volumes  based on pairwise comparisons: Bees fed Rockrose pollen had significantly elevated IAPV titer volumes with virus infection compared to uninfected controls (Tukey HSD, p $<$ 5.44e-03). However, bees fed Chestnut pollen did not have significantly elevated IAPV titer volumes with virus infection compared to uninfected controls (Tukey HSD, p = 1.11e-01). Overall, we interpreted these findings to mean that high-quality Chestnut pollen could ``rescue'' high virus titers resulting from the inoculation treatment, whereas low-quality Rockrose pollen could not do so (Figure 2).

@@

\section*{Ethics approval and consent to participate}
  Not applicable.

>>>>

As shown in Figure 1, mortality rates of honey bees 72 hour post-inoculation significantly differed among the treatment groups (mixed model ANOVA across all treatment groups, df = 3, 54; F = 10.03; p $<$ 2.30e-05). The effect of virus treatment (mixed model ANOVA, df = 1, 54; F = 24.73; p $<$ 1.00e-05) and diet treatment (mixed model ANOVA, df = 1, 54; F = 5.32; p $<$ 2.49e-02) were significant, but the interaction between the two factors (mixed model ANOVA, df = 1, 54; F = 4.72e-02, p = 8.29e-01) was not significant. We compared mortality levels based on pairwise comparisons: For a given diet, honey bees exposed to the virus showed significantly higher mortality rate than honey bees not exposed to the virus. Namely, bees fed Rockrose pollen had significantly elevated mortality with virus infection compared to uninfected controls (Tukey HSD, p $<$ 1.31e-03), and bees fed Chestnut pollen similarly had significantly elevated mortality with virus infection compared to controls (Tukey HSD, p $<$ 4.78e-03) (Figure 1).

As shown in Figure 2, IAPV titers of honey bees 72 hour post-inoculation significantly differed among the treatment groups (mixed model ANOVA across all treatment groups, df = 3, 33; F = 6.10; p $<$ 1.96e-03). The effect of virus treatment (mixed model ANOVA, df = 1, 33; F = 15.04; p $<$ 4.75e-04) was significant, but the diet treatment (mixed model ANOVA, df = 1, 33; F = 2.55; p = 1.20e-01) and the interaction between the two factors (mixed model ANOVA, df = 1, 33; F = 7.02e-01, p = 4.08e-01) were not significant. We compared IAPV titer volumes  based on pairwise comparisons: Bees fed Rockrose pollen had significantly elevated IAPV titer volumes with virus infection compared to uninfected controls (Tukey HSD, p $<$ 5.65e-03). However, bees fed Chestnut pollen did not have significantly elevated IAPV titer volumes with virus infection compared to uninfected controls (Tukey HSD, p = 1.11e-01). Overall, we interpreted these findings to mean that high-quality Chestnut pollen could ``rescue'' high virus titers resulting from the inoculation treatment, whereas low-quality Rockrose pollen could not do so (Figure 2).

@@

\section*{Ethics approval and consent to participate}
  All honey bees used in this work were sampled in the United States, and no ethical use approval is required for this species in this country.
